---
title: "ST558 Project 1"
author: "Yu Jiang"
date: "6/10/2020"
output: 
  rmarkdown::github_document
---
# 1. Describe JSON Data

## What is JSON?

* Json stands for JavaScript Object Notation.

* The lightweight format for storing and transporting data, specified by Douglas Crockford.

* A way to store information in an organized, easy-to-access manner. 

* It has been extended from the JavaScript scripting language.

* It was designed for human-readable data interchange.


## Where does it get used? 

* It is used when data is sent from a server to a web page. 

* The JSON format is used for serializing and transmitting structured data over network connection.

* It is used while writing JavaScript based applications that includes browser extensions and websites. 

* Web services and APIs use JSON format to provide public data. 

* It can be used with modern programming languages.

## Why is it a good way to store data?

* It is 'self-describing' and easy to understand.

* It is a generic data format with a minimal number of value types: strings, numbers, booleans, lists, objects, and null.

* It can be supported by many databases like PostgreSQL and MySQL.

* JSON data is stored in a set of key-value pairs.

* Spacing (spaces, tabs, new lines) does not matter in a JSON file.

For more information about JSON, please check [this website](https://en.wikipedia.org/wiki/JSON). 

# 2. Three Packages for Reading JSON Data into R

## 'rjson'

The package 'rjson' converts a JSON object into an R object. 

## 'RJSONIO'

The package 'RJSONIO' allows conversion to and from
data in Javascript object notation (JSON) format.
This allows R objects to be inserted into Javascript/ECMAScript/ActionScript code
and allows R programmers to read and convert JSON content to R objects.

## 'jsonlite'

A fast JSON parser and generator optimized for statistical data and the web.

I am going to choose the package 'jsonlite' since it is the fastest among these three packages and this package offers flexible, robust, high performance tools for working with JSON in R and is particularly powerful for building pipelines and interacting with a web API. 

# 3. Write Functions to Contact the NHL Records

```{r, message=FALSE}
# Load the libraries
library(httr)
library(jsonlite)
library(dplyr)

# The full_url for the first call

full_url <- 'https://records.nhl.com/site/api/franchise'

# Use a GET command to retrieve the franchise information in raw form
franchise <- GET(full_url)
# franchise

# Use 'content' to turn it into JSON text form
frc_text <- content(franchise, "text")
# Use 'fromJSON' with 'flatten = TRUE' to convert it to a list
frc_json <- fromJSON(frc_text, flatten = TRUE)

# Access the data
attributes(frc_json)
frc_json$data %>% head()
frc_json$total


```

```{r}
# Start with a base_URL and paste the pieces together

full_url <- 'https://records.nhl.com/site/api/franchise-team-totals'

# Use a GET command to retrieve the franchise information in raw form
franchise_team <- GET(full_url)
franchise_team

# Use 'content' to turn it into JSON text form
frc_team_text <- content(franchise_team, "text")
# Use 'fromJSON' with 'flatten = TRUE' to convert it to a list
frc_team_json <- fromJSON(frc_team_text, flatten = TRUE)

# Access the data
attributes(frc_team_json)
frc_team_json$data
frc_team_json$total

```


```{r}
# Start with a base_URL and paste the pieces together

full_url <- 
  'https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=12'

# Use a GET command to retrieve the franchise information in raw form
franchise_team <- GET(full_url)
franchise_team

# Use 'content' to turn it into JSON text form
frc_team_text <- content(franchise_team, "text")
# Use 'fromJSON' with 'flatten = TRUE' to convert it to a list
frc_team_json <- fromJSON(frc_team_text, flatten = TRUE)

# Access the data
attributes(frc_team_json)
frc_team_json$data
frc_team_json$total
```

```{r}
# Start with a base_URL and paste the pieces together

full_url <- 
  'https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=10'

# Use a GET command to retrieve the franchise information in raw form
franchise_team <- GET(full_url)
franchise_team

# Use 'content' to turn it into JSON text form
frc_team_text <- content(franchise_team, "text")
# Use 'fromJSON' with 'flatten = TRUE' to convert it to a list
frc_team_json <- fromJSON(frc_team_text, flatten = TRUE)

# Access the data
attributes(frc_team_json)
frc_team_json$data
frc_team_json$total
```

```{r}
# Start with a base_URL and paste the pieces together

full_url <- 
  'https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=10'

# Use a GET command to retrieve the franchise information in raw form
franchise_team <- GET(full_url)
franchise_team

# Use 'content' to turn it into JSON text form
frc_team_text <- content(franchise_team, "text")
# Use 'fromJSON' with 'flatten = TRUE' to convert it to a list
frc_team_json <- fromJSON(frc_team_text, flatten = TRUE)

# Access the data
attributes(frc_team_json)
frc_team_json$data
frc_team_json$total
```



# 4. Data Analysis

```{r, messagae=FALSE}
# Load the libraries
library(DT)
library(knitr)
library(tidyverse)
library(ggplot2)
```

