---
title: "ST558 Project 1"
author: "Yu Jiang"
date: "6/10/2020"
output: 
  rmarkdown::github_document:
    toc: yes
---
# 1. Describe JSON Data

## What is JSON?

* Json stands for JavaScript Object Notation.

* The lightweight format for storing and transporting data, specified by Douglas Crockford.

* A way to store information in an organized, easy-to-access manner. 

* It has been extended from the JavaScript scripting language.

* It was designed for human-readable data interchange.


## Where does it get used? 

* It is used when data is sent from a server to a web page. 

* The JSON format is used for serializing and transmitting structured data over network connection.

* It is used while writing JavaScript based applications that includes browser extensions and websites. 

* Web services and APIs use JSON format to provide public data. 

* It can be used with modern programming languages.

## Why is it a good way to store data?

* It is 'self-describing' and easy to understand.

* It is a generic data format with a minimal number of value types: strings, numbers, booleans, lists, objects, and null.

* It can be supported by many databases like PostgreSQL and MySQL.

* JSON data is stored in a set of key-value pairs.

* Spacing (spaces, tabs, new lines) does not matter in a JSON file.

For more information about JSON, please check [this website](https://en.wikipedia.org/wiki/JSON). 

# 2. Three Packages for Reading JSON Data into R

## 'rjson'

The package 'rjson' converts a JSON object into an R object. 

## 'RJSONIO'

The package 'RJSONIO' allows conversion to and from
data in Javascript object notation (JSON) format.
This allows R objects to be inserted into Javascript/ECMAScript/ActionScript code
and allows R programmers to read and convert JSON content to R objects.

## 'jsonlite'

A fast JSON parser and generator optimized for statistical data and the web.

I am going to choose the package 'jsonlite' since it is the fastest among these three packages and this package offers flexible, robust, high performance tools for working with JSON in R and is particularly powerful for building pipelines and interacting with a web API. 

# 3. Write Functions to Contact the NHL Records

```{r, message=FALSE}

# Load the libraries
library(httr)
library(jsonlite)
library(dplyr)

# A function for the first two calls
base_url_1 <- 'https://records.nhl.com/site/api'

api <- function(path){
 json_1 <- GET(paste0(base_url_1, "/", path)) %>%
    content('text') %>%
    fromJSON(flatten = TRUE) 
 
 attributes(json_1)
 return(tbl_df(json_1$data))
}

# Table for /frachise
api('franchise') %>% head()

# Table for /franchise-team-totals
api('franchise-team-totals') %>% head()

# A function for the last three calls
base_url_2 <- 'https://records.nhl.com/site/api/franchise'

ID <- function(arg1, arg2){
 json_2 <- GET(paste0(
    base_url_2, "-", arg1, '-records?cayenneExp=franchiseId=', arg2)) %>%
    content('text') %>%
    fromJSON(flatten = TRUE)
 
 attributes(json_2)
 return(tbl_df(json_2$data))
}

# Table for /franchise-season-records?cayenneExp=franchiseId=ID
ID('season', '12') %>% head()

# Table for /franchise-goalie-records?cayenneExp=franchiseId=ID
ID('goalie', '12') %>% head()

# Table for //franchise-skater-records?cayenneExp=franchiseId=ID
ID('skater', '12') %>% head()

```


# 4. Data Analysis

```{r, message=FALSE}
# Load the libraries
library(tidyverse)
library(knitr)
library(ggplot2)

# Create some new variables, the percent of home wins/losses and road wins/losses
team_total <- api('franchise-team-totals') %>%
  select(
    -c('id', 'firstSeasonId', 'penaltyMinutes', 'pointPctg', 'points')) %>% 
  mutate(HomeWinsPercent = homeWins/wins, 
         HomeLossesPercent = homeLosses/losses, 
         RoadWinsPercent = roadWins/wins, 
         RoadLossesPercent = roadLosses/losses)


# Convert activeFranchise and gameTypeId to factors and rename their with levels
team_total$activeFranchise <- as.factor(team_total$activeFranchise)
levels(team_total$activeFranchise) <- list('Active'= 1, 'Non-Active'= 0)

team_total$gameTypeId <- as.factor(team_total$gameTypeId)
levels(team_total$gameTypeId) <- list('Type 2'= 2, 'Type 3'= 3)

# Select the rows of gameType ID = 2 and gameType ID = 3
id_2 <- team_total[team_total$gameTypeId == 'Type 2', ]
id_3 <- team_total[team_total$gameTypeId == 'Type 3', ]

# Table about the preview of team total data
knitr::kable(
 head(team_total),
  caption = 'Preview of Team Total Data')

# Table about the ActiveFranchise vs Game Types
knitr::kable(
 table(team_total$activeFranchise, team_total$gameTypeId),
  caption = 'ActiveFranchise vs Game Types')

# Table about the wins data for game type = 2
wins2_a <- id_2 %>% select(activeFranchise,wins, homeWins, roadWins) %>%
  filter(activeFranchise == 'Active') %>% 
  summarise(
    Wins = sum(wins), HomeWins = sum(homeWins), RoadWins = sum(roadWins))

wins2_non <- id_2 %>% select(activeFranchise,wins, homeWins, roadWins) %>%
  filter(activeFranchise == 'Non-Active') %>% 
  summarise(
    Wins = sum(wins), HomeWins = sum(homeWins), RoadWins = sum(roadWins))

table_type_2_win <- data.frame(
  Status = c('Active', 'Non-Active'), Wins = c(50938, 1147), HomeWins = c(29436, 718), RoadWins = c(21502, 429))

knitr::kable(
  (table_type_2_win),
  caption = 'Detailed Wins for Game Type 2 '
  )

# Table about the loss data for game type = 2
loss2_a <- id_2 %>% select(activeFranchise,losses, homeLosses, roadLosses) %>%
  filter(activeFranchise == 'Active') %>% 
  summarise(
    Losses = sum(losses), HomeLosses = sum(homeLosses), RoadLosses = sum(roadLosses))

loss2_non <- id_2 %>% 
  select(activeFranchise,losses, homeLosses, roadLosses) %>%
  filter(activeFranchise == 'Non-Active') %>% 
  summarise(
    Losses = sum(losses), HomeLosses = sum(homeLosses), RoadLosses = sum(roadLosses))

table_type_2_loss <- data.frame(
  Status = c('Active', 'Non-Active'), Losses = c(45595, 1663), HomeLosses = c(18925, 660), RoadLosses = c(26670, 1003))

knitr::kable(
  (table_type_2_loss),
  caption = 'Detailed Losses for Game Type 2 '
  )

```

```{r}
# Table about the wins data for game type = 3
wins3_a <- id_3 %>% select(activeFranchise,wins, homeWins, roadWins) %>%
  filter(activeFranchise == 'Active') %>% 
  summarise(
    Wins = sum(wins), HomeWins = sum(homeWins), RoadWins = sum(roadWins))

wins3_non <- id_3 %>% select(activeFranchise,wins, homeWins, roadWins) %>%
  filter(activeFranchise == 'Non-Active') %>% 
  summarise(
    Wins = sum(wins), HomeWins = sum(homeWins), RoadWins = sum(roadWins))

table_type_3_win <- data.frame(
  Status = c('Active', 'Non-Active'), Wins = c(4375, 48), HomeWins = c(2505, 21), RoadWins = c(1870, 27))

knitr::kable(
  (table_type_3_win),
  caption = 'Detailed Wins for Game Type 3 '
  )

# Table about the loss data for game type = 3
loss3_a <- id_3 %>% select(activeFranchise,losses, homeLosses, roadLosses) %>%
  filter(activeFranchise == 'Active') %>% 
  summarise(
    Losses = sum(losses), HomeLosses = sum(homeLosses), RoadLosses = sum(roadLosses))

loss3_non <- id_3 %>% 
  select(activeFranchise,losses, homeLosses, roadLosses) %>%
  filter(activeFranchise == 'Non-Active') %>% 
  summarise(
    Losses = sum(losses), HomeLosses = sum(homeLosses), RoadLosses = sum(roadLosses))

table_type_3_loss <- data.frame(
  Status = c('Active', 'Non-Active'), Losses = c(4352, 59), HomeLosses = c(1866, 27), RoadLosses = c(2486, 32))

knitr::kable(
  (table_type_3_loss),
  caption = 'Detailed Losses for Game Type 3 '
  )

```

```{r}
# 

